---
title: "Getting Started with seroreconstruct"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with seroreconstruct}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Overview

`seroreconstruct` is an R package for Bayesian inference of influenza virus
infection status from serial antibody measurements. It relaxes the traditional
4-fold rise rule by jointly modeling antibody boosting after infection, waning
over time, and measurement error — estimating individual-level infection
probabilities from hemagglutination inhibition (HAI) titer data.

The statistical framework is described in
[Tsang et al. (2022) *Nature Communications* 13:1557](https://doi.org/10.1038/s41467-022-29233-8).

## Input data

The package requires two data frames:

1. **`inputdata`** — one row per individual with columns:
   - `age_group`: integer (0 = children, 1 = adults, 2 = older adults)
   - `start_time`, `end_time`: follow-up period (integer days)
   - `time1`, `time2`, `time3`: serum collection dates (integer days)
   - `HAI_titer_1`, `HAI_titer_2`, `HAI_titer3`: HAI titers at each collection

2. **`inputILI`** — daily influenza-like illness activity, with rows corresponding
   to consecutive days. The row indices must cover the range of `start_time` to
   `end_time` in `inputdata`.

```{r load-data}
library(seroreconstruct)

data("inputdata")
data("flu_activity")

head(inputdata)
dim(inputdata)
```

## Fitting the model

The main function `sero_reconstruct()` runs a Bayesian MCMC to estimate infection
probabilities and antibody dynamics parameters. For real analyses, use at least
100,000 iterations with appropriate burn-in and thinning.

```{r fit-model, eval = FALSE}
fit <- sero_reconstruct(inputdata, flu_activity,
                        n_iteration = 200000, burnin = 100000, thinning = 10)
```

For this vignette, we use a short run for illustration:

```{r fit-short}
fit <- sero_reconstruct(inputdata, flu_activity,
                        n_iteration = 5000, burnin = 2000, thinning = 1)
fit
```

## Viewing results

The `summary()` method extracts key estimates with 95% credible intervals:

```{r summary}
summary(fit)
```

The summary table includes:

- **Random error (%)** and **Two-fold error (%)**: measurement error parameters
- **Boosting**: fold-increase in antibody titer after infection
- **Waning**: fold-decrease in antibody titer per year post-infection
- **Infection probability**: overall and stratified by baseline HAI titer

## Visualization

### MCMC diagnostics

Check convergence with trace plots and posterior density plots:

```{r diagnostics, fig.height = 8}
plot_diagnostics(fit, params = c("random_error", "twofold_error"))
```

### Antibody trajectories

Visualize posterior trajectories for individual participants. Red lines show
trajectories where infection occurred; blue lines show trajectories without
infection.

```{r trajectory, fig.height = 4}
plot_trajectory(fit, id = 1)
```

### Boosting and waning

```{r boosting-waning, fig.height = 4, fig.width = 9}
par(mfrow = c(1, 2))
plot_boosting(fit)
plot_waning(fit)
```

### Infection probabilities

Forest plot of posterior infection probabilities with 95% credible intervals:

```{r infection-prob, fig.height = 3}
plot_infection_prob(fit)
```

## Tables

Extract parameter estimates and individual-level results as data frames:

```{r tables}
# Model parameter estimates
table_parameters(fit)

# Per-individual infection probabilities
head(table_infections(fit))
```

## Subgroup analysis

Compare infection rates across groups by fitting independent MCMCs:

```{r group-by, eval = FALSE}
fit_by_age <- sero_reconstruct(inputdata, flu_activity,
                               n_iteration = 200000, burnin = 100000,
                               thinning = 10, group_by = ~age_group)

summary(fit_by_age)
```

## Joint model with shared parameters

When comparing groups, measurement error is a lab assay property — it is shared
across all groups. You can additionally share boosting/waning if the groups are
expected to have similar antibody dynamics:

```{r shared, eval = FALSE}
# Share all parameters except infection probability
fit_joint <- sero_reconstruct(inputdata, flu_activity,
                              n_iteration = 200000, burnin = 100000,
                              thinning = 10,
                              group_by = ~age_group,
                              shared = c("error", "boosting_waning"))

summary(fit_joint)
```

## Simulation

Generate synthetic data for power analysis or validation:

```{r simulate}
data("para1")
data("para2")

sim <- simulate_data(inputdata, flu_activity, para1, para2)
names(sim)
```

The simulated data can be passed directly to `sero_reconstruct()` for
simulation-recovery studies.

## Citation

To cite seroreconstruct in publications:

> Tsang TK, Perera RAPM, Fang VJ, Wong JY, Shiu EY, So HC, Ip DKM,
> Malik Peiris JS, Leung GM, Cowling BJ, Cauchemez S. (2022).
> Reconstructing antibody dynamics to estimate the risk of influenza virus
> infection. *Nature Communications* 13:1557.
> doi: [10.1038/s41467-022-29233-8](https://doi.org/10.1038/s41467-022-29233-8)
